Lucy.js: Client-Side Indexes for Fast Full-Text Searching

### 2014-12-11 paper/paper.tex
cleaned up some of the section on tries

Many applications on the web today use a combination of client-side and server-side data stores to facilitate fast interactive and data-intensive experiences. 
However, standard client side databases within browsers do not currently support full-text searching, and it is not clear the best way to engineer full-text indexes on top of the existing infrastructure.
In this paper, we introduce Lucy.js, a fully client-side search engine built as an extension of the existing well-supported IndexedDB API.
We compare the performance of the different indexes we implement on different types of full-text content and queries. We find that a hybrid approach combining inverted indexes and tries perform the best, particularly on larger datasets.
We also implement various real-life applications of our system and compare the performance of our system with that of using a fully server side system. We find that given the costs of network latency, our approach is signficantly faster than relying on a server-side search engine.
Finally, we discuss how our system may open up novel web application and data storage designs that solely use client-side search engines or combine both client and server search engines.
\end{abstract}

@@

One function that we provide is \texttt{Lucy.calculateWeight()}, which counts the number of words in the search query that appear in the document. The other ranking we provide is \texttt{Lucy.calculateCoverDensity()}, which implements Cover Density Ranking~\cite{li2002improvement}, an algorithm that takes into account positional information. For this algorithm to work, we need to store at index creation time the positions that the word appears in within the document. Both functions take in an optional \texttt{normalization} integer where the developer can specify whether they are interested in normalizing by document length or unique number of words in the document. If normalization is desired, then at index creation time, we calculate these measures and store them with the document object.

@@

A trie index in Lucy.js (prefix or suffix trie) is represented by two underlying entities in IndexedDB: an objects store to save nodes in the trie, and an index on that object store to make trie traversal faster.  A reference to the input object store to be indexed -- which contains the full-text to be searched -- is also saved.

\subsubsection{Trie Node Object Store}
Each entry in the object store represents a different node in the trie. The structure of entries in the object store is as follows:
\begin{center}
node\_id: \{ parent\_id: a, char: b, doc\_ids: [x,y,z] \}
\end{center}
The key to the object store is the node\_id, and the corresponding value object has three attributes: parent\_id, char, and doc\_ids. The parent\_id attribute is the node id of the parent node in the trie, and the char attribute is the character value for the trie node. The final attribute is doc\_ids, which is a list of keys into the input text object store. The keys in this list are like foreign keys that allow the trie index to retrieve the actual text from the input object store when answering queries. Each path -- and therefore each node -- in the trie represents a unique prefix/suffix. The doc\_ids for a node contains the ids of all documents with that given prefix/suffix. Inserting a word from a given document therefore involves appending that document's id into the doc\_ids of each node along the path corresponding to the word. The structure of the trie is shown in Figure ******.  It is important to note that this approach is not very scalable; since each node stores a list of matching document ids, interior nodes closer to the root of the trie could store large lists of these ids. However, the benefit is that searches in the trie do not require traversal all the way down to the leaves. Matches can be returned as soon as the prefix is traced in the trie, and no further traversal along subtrees is needed. An alternative design to the trie index is the Hybrid Trie discussed in the next section.

\begin{figure}[h!]
  \centering
   \includegraphics[scale=.35]{trie_figure}
  \caption{Structure of a trie index with two input tweets. The tweets are first tokenized and stop words are removed. Note that document ids are stored in every node. An artificial root node is created as the root of the trie.}
\end{figure}

\subsubsection{Trie Node Index}
While the structure of the trie node object store optimizes lookups by node id, this is not particularly useful for trie traversal. When traversing down a trie, we would instead like to look up nodes by both their character value and their parent id. It would have been possible to change the design of the trie object store so that each node stored a list of children node ids instead of a single parent id. However, this approach requires each node to store more information, since a list of children ids is saved per node instead of a single parent id. Additionally, we would then need to iterate over this list of children ids in order to find a particular child node with a given character value. With the current object store structure, we can optimize lookups by parent\_id and char by building an index on both these attributes. Overall, the trie index is backed by an object store for node storage and an index on that object store with a compound key of both parent\_id and char.

\subsubsection{Lookups}
Lookups in the trie index are performed by traversing the trie downward, character by character in the search pattern. Partial matches are not accepted; the entire prefix (or suffix) must be present in the trie or no results are returned. If the entire pattern is found, the doc\_ids are extracted from the trie node corresponding to the last letter of the search pattern. These ids are used to key into the input text object store and retrieve the original text data.

@@

\subsubsection{Depth-Limiting}
To limit the size of trie indexes, an optional maximum depth can be set. If set, words longer than the depth are cut off so that the remainder substrings are stored at the leaves. Figure ****** shows the structure of the same trie as before, but with a maximum depth of 4. Leaf nodes store this substring data as a mapping from substring to document ids containing that substring. During lookups, if the maximum depth of the trie is reached, this map is searched for matches. So while depth-limiting trie indexes can limit the number of trie nodes that need to be traversed upon lookups, they incur the cost of searching these maps at the leaves. If the max depth parameter is too small, these maps can be large and slow down search significantly. If set too large, the max depth parameter has no effect.

\begin{figure}[h!]
  \centering
   \includegraphics[scale=.35]{trie_maxdepth_figure}
  \caption{Structure of a trie index with a maximum depth of 4. If a word gets cut off by the depth limit, the substring remainder is stored at the leaves. Each substring remainder maps to a list of document ids with that remainder.}
\end{figure}

>>>>

Many applications on the web today use a combination of client-side and server-side data stores to facilitate fast interactive and data-intensive experiences. 
However, standard client side databases within browsers do not currently support full-text searching, and it is not clear the best way to engineer full-text indexes on top of the existing infrastructure.
In this paper, we introduce Lucy.js, a fully client-side search engine built as an extension of the existing well-supported IndexedDB API.
We compare the performance of the different indexes we implement on different types of full-text content and queries. We find that a hybrid approach combining inverted indexes and tries perform the best, particularly on larger datasets.
We also implement various real-life applications of our system and compare the performance of our system with that of using a fully server side system. We find that given the costs of network latency, our approach is significantly faster than relying on a server-side search engine.
Finally, we discuss how our system may open up novel web application and data storage designs that solely use client-side search engines or combine both client and server search engines.
\end{abstract}

@@

One function that we provide is \texttt{Lucy.calculateWeight()}, which counts the number of words in the search query that appear in the document. The other ranking we provide is \texttt{Lucy.calculateCoverDensity()}, which implements Cover Density Ranking~\cite{li2002improvement}, an algorithm that takes into account positional information. The result is that documents that contain the terms in the search query closer together are ranked higher than documents that have them farther apart. For this algorithm to work, we need to store at index creation time the positions that the word appears in within the document.

Both functions take in an optional \texttt{normalization} integer where the developer can specify whether they are interested in normalizing by document length or unique number of words in the document. If normalization is desired, then at index creation time, we calculate these measures and store them with the document object. Our current implementation stores them in the original object store, but another implementation could store them in a separate object store for document metadata.

@@

A trie index in Lucy.js (prefix or suffix trie) is represented by two underlying entities in IndexedDB: an object store to save nodes in the trie, and an index on that object store to make trie traversal faster.  A reference to the input object store to be indexed, which contains the full-text to be searched, is also saved.

\subsubsection{Trie Node Object Store}
Each entry in the object store represents a different node in the trie. The structure of entries in the object store is as follows:
\begin{center}
\texttt{node\_id:\{parent\_id:a, char:b, docIds:[x,y,z]\}}
\end{center}
The key to the object store is the \texttt{node\_id}, and the corresponding value object has three attributes: \texttt{parent\_id}, \texttt{char}, and \texttt{docIds}. The \texttt{parent\_id} attribute is the node id of the parent node in the trie, and the \texttt{char} attribute is the character value for the trie node. The final attribute is \texttt{doc\_ids}, which is a list of keys into the input text object store. The keys in this list are like foreign keys that allow the trie index to retrieve the actual text from the input object store when answering queries. Each path, and therefore each node, in the trie represents a unique prefix/suffix. The \texttt{docIds} for a node contains the ids of all documents with that given prefix/suffix. Inserting a word from a given document therefore involves appending that document's id into the \texttt{docIds} of each node along the path corresponding to the word. The structure of the trie is shown in Figure~\ref{trie}.  It is important to note that this approach is not very scalable; since each node stores a list of matching document ids, interior nodes closer to the root of the trie could store large lists of these ids. However, the benefit is that searches in the trie do not require traversal all the way down to the leaves. Matches can be returned as soon as the prefix is traced in the trie, and no further traversal along subtrees is needed. This is useful for when the dataset in question is reasonably small and speed is of great importance. An alternative design to the trie index is the Hybrid Trie discussed in the next section.

\begin{figure}[h!]
  \centering
   \includegraphics[scale=.35]{trie_figure}
  \caption{Structure of a trie index with two input tweets. The tweets are first tokenized and stop words are removed. Note that document ids are stored in every node. An artificial root node is created as the root of the trie.}
\label{trie}
\end{figure}

\subsubsection{Trie Node Index}
While the structure of the trie node object store optimizes lookups by node id, this is not particularly useful for trie traversal. When traversing down a trie, we would instead like to look up nodes by both their character value and their parent id. It would have been possible to change the design of the trie object store so that each node stored a list of children node ids instead of a single parent id. However, this approach requires each node to store more information, since a list of children ids is saved per node instead of a single parent id. Additionally, we would then need to iterate over this list of children ids in order to find a particular child node with a given character value. With the current object store structure, we can optimize lookups by \texttt{parent\_id} and \texttt{char} by building an index on both these attributes. Overall, the trie index is backed by an object store for node storage and an index on that object store with a compound key of both \texttt{parent\_id} and \texttt{char}.

\subsubsection{Lookups}
Lookups in the trie index are performed by traversing the trie downward, character by character in the search pattern. Partial matches are not accepted; the entire prefix (or suffix) must be present in the trie or no results are returned. If the entire pattern is found, the \texttt{docIds} are extracted from the trie node corresponding to the last letter of the search pattern. These ids are used to key into the input text object store and retrieve the original text data.

@@

\subsubsection{Depth-Limiting}
To limit the size of trie indexes, an optional maximum depth can be set. If set, words longer than the depth are cut off so that the remainder substrings are stored at the leaves. Figure~\ref{triemax} shows the structure of the same trie as before, but with a maximum depth of 4. Leaf nodes store this substring data as a mapping from substring to document ids containing that substring. During lookups, if the maximum depth of the trie is reached, this map is searched for matches. So while depth-limiting trie indexes can limit the number of trie nodes that need to be traversed upon lookups, they incur the cost of searching these maps at the leaves. If the max depth parameter is too small, these maps can be large and slow down search significantly. If set too large, the max depth parameter has no effect.

\begin{figure}[h!]
  \centering
   \includegraphics[scale=.35]{trie_maxdepth_figure}
  \caption{Structure of a trie index with a maximum depth of 4. If a word gets cut off by the depth limit, the substring remainder is stored at the leaves. Each substring remainder maps to a list of document ids with that remainder.}
\label{triemax}
\end{figure}
